---
title: "Lab 1: Intro to R"
author: "Daniel Craig"
date: "`r Sys.Date()`"
output: openintro::lab_report
---

```{r load-packages, message=FALSE}
library(tidyverse)
library(openintro)
```

### Exercise 1

```{r view-girls-counts}
arbuthnot$girls
```


### Exercise 2

There seems to be a significant increase in girls baptized per year starting at 1660. I assume some type of religious movement sparked at that time.

```{r trend-girls}
ggplot(data = arbuthnot, aes(x = year, y = girls)) + 
  geom_point()
```


### Exercise 3

It doesn't seem like a strong relationship but a downward trend from 1660 onward seems to be present.

```{r plot-prop-boys-arbuthnot}
arbuthnot <- arbuthnot %>%
  mutate(total = boys + girls)

arbuthnot <- arbuthnot %>%
  mutate(boy_to_girl_ratio = boys / girls)

arbuthnot <- arbuthnot %>%
  mutate(boy_ratio = boys / total)

ggplot(data = arbuthnot, aes(x = year, y = boy_ratio)) + 
  geom_point()
```


### Exercise 4
  
Years included in the data are 1940 - 2002  
  
Dimensions are 63 rows by 3 columns  
  
After our manipulation, names of the columns are: year,boys,girls
  

```{r dim-present}
data('present',package = 'openintro')
unique(present$year)
dim(present)
names(present)
```


### Exercise 5

Not surprisingly, modern population counts are of magnitudes higher. Although their average ratios are similar. Present day US has a slightly lower boy to girl ratio of 1.05 compared to arbuthnot's 1.07

```{r count-compare}
arbuthnot %>%
  summarize(min = min(boys), max = max(boys))

present %>%
  summarize(min = min(boys), max = max(boys))

present <- present %>%
  mutate(total = boys + girls)

present <- present %>%
  mutate(boy_to_girl_ratio = boys / girls)

present <- present %>%
  mutate(boy_ratio = boys / total)

mean(present$boy_to_girl_ratio)
mean(present$boy_ratio)
mean(arbuthnot$boy_to_girl_ratio)
mean(arbuthnot$boy_ratio)

```


### Exercise 6

The observation holds, but is decreasing and slightly less than arbuthnot's findings.

```{r plot-prop-boys-present}
ggplot(data = present, aes(x = year, y = boy_ratio)) + 
  geom_point()
```


### Exercise 7

1961 had the highest total number of births.

```{r find-max-total}
present <- present %>%
  mutate(total = boys + girls)

present %>% arrange(desc(total))
```

